---
title: "Regime Changes and Economic Preferences: Global Evidence"
subtitle: "Milestone 3: Econometric Approach"
author: "Andrea Češková and Elvin Mammadov"
format: 
  pdf:
    reference-location: document
number-sections: true
toc: true
crossref:
  fig-title: "Figure"
  tbl-title: "Table"
  eq-title: "Equation"
  lot-title: "List of figures"
bibliography: references.bib
link-citations: true
---

```{r}
#| message: false
#| warning: false
#| include: false
#  Load the function from extract_codes.R

source("../Input/extract_codes.R")

#  Use it to extract code from the Rmd into an R script
extract_r_code_from_rmd("../Input/Data loading + wrangling.Rmd", "../Input/data_wrangle.R")

#  Run the extracted R code
source("../Input/data_wrangle.R")

write_dta(final_data_gdp, path = "../Input/final_data_gdp.dta")

```

# Economic model

As Detlefsen et al. [-@detlefsen_are_2024] points out, the formation of economic preferences in **childhood** and **adolescence** has a long-term consequences for life outcomes. To provide an example, patience can impact one's school performance, which can later translate to education attainment and labor market outcomes. The same way can risk preference impact one's financial decisions [@detlefsen_are_2024].

According to the National Library of Medicine, childhood is defined as a stage from **3-11 years** and adolescence as a stage from **12-18 years**[@balasundaram2025]. But what factors can have an impact on economic preferences?

# Setting

## Treatment Definition

Treatment definition was one of our main tasks until now. Our goal was to capture both sudden and gradual regime changes, as well as the direction of change (democratization or autocratization), based on changes in the Liberal Democracy Index (LDI).

We followed a methodology similar to that described in [@luhrmann_autocratization_2020], which identifies regime changes by evaluating differences in the LDI between year $t$ and year $t-10$. However, while the original study used a threshold of **0.05**, we adopted a stricter threshold of **0.1**. This modification was made intentionally to focus on more **sudden and substantial regime shifts**, rather than more gradual changes.

The process includes computing standard errors and confidence intervals of LDI changes. A regime change is considered statistically significant if the confidence intervals of the LDI values do not overlap and the absolute change in LDI exceeds our chosen threshold of 0.1:

$$
|\Delta \text{LDI}_{t}| = |\text{LDI}_{t} - \text{LDI}_{t-10}| > 0.1
$$

Once a regime change is identified for a specific country and year, we link this information to the GPS dataset and assign treatment to individuals who were between the ages of 3 and 18 during the year of regime change in their respective countries.

```{r}
#| echo: false
#| label: tbl-c
#| tbl-cap: Number of treatment variables
#| tbl-pos: "H"
table(final_data$formative_regime_change)
```

```{r}
#| echo: false
#| label: fig-a
#| warning: false
#| results: 'asis'
#| fig-cap: "Mean Preference Profiles by Regime Change Exposure"
#| fig-cap-location: top
#| fig-pos: "H"
p_load(dplyr,
       kableExtra)
long_data <- final_data_gdp %>%
  pivot_longer(
    cols = c(trust, patience, risktaking, posrecip, negrecip, altruism),
    names_to = "preference_type",
    values_to = "value"
  ) %>%
  mutate(preference_type = str_to_title(preference_type))

# Change the formative_regime_change factor levels and labels
long_data <- long_data %>%
  mutate(formative_regime_change = case_when(
    formative_regime_change == 0 ~ "Control",
    formative_regime_change == 1 ~ "Treated",
    TRUE ~ as.character(formative_regime_change)
  ))

# Calculating group means and confidence intervals
group_means <- long_data %>%
  group_by(formative_regime_change, preference_type) %>%
  summarise(
    mean_value = mean(value, na.rm = TRUE),
    se = sd(value, na.rm = TRUE) / sqrt(n()),
    ci_lower = mean_value - 1.96 * se,
    ci_upper = mean_value + 1.96 * se,
    .groups = "drop"
  )

# Order factors to make Treated come FIRST so it gets the first color (red in Set1)
group_means$formative_regime_change <- factor(
  group_means$formative_regime_change,
  levels = c("Treated", "Control")
)

# Ordering from smallest to highest differences between groups
differences <- group_means %>%
  select(preference_type, formative_regime_change, mean_value) %>%
  pivot_wider(names_from = formative_regime_change, values_from = mean_value) %>%
  mutate(difference = abs(Treated - Control)) %>%
  arrange(difference) %>%
  select(preference_type, difference)

# Create ordered factor for preference_type based on differences
group_means$preference_type <- factor(group_means$preference_type, 
                                      levels = differences$preference_type)

grouped_bar_plot <- ggplot(group_means, aes(x = preference_type, y = mean_value, fill = formative_regime_change)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  geom_text(aes(label = round(mean_value, 3)), 
            position = position_dodge(width = 0.8), 
            vjust = ifelse(group_means$mean_value >= 0, -0.3, 1.2),
            size = 3.5,
            fontface = "bold") +
  scale_fill_manual(values = c("Treated" = "#D62728", "Control" = "#1F77B4"),
                    name = "Group") +
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.1))) +  # Add space for labels
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_text(size = 12),
    legend.position = "top"
  ) +
  labs(
    caption = "Mean values displayed above/below bars",
    x = element_blank(),
    y = "Mean Value"
  ) +
  geom_hline(yintercept = 0, linetype = "dotted", color = "gray50", size = 0.5)

# Print the plot
print(grouped_bar_plot)
```

In the table @tbl-preferences we examine the difference in mean in mean on a more formal level. We perform t-tests of the differences. All of them seems to be statistically significant on a 0.001 confidence level. This points out in a direction for our analysis.

```{r}
#| include: false
# Function to perform t-tests and format results
create_ttest_table <- function(data, vars) {
  
  # Initialize results list
  results <- list()
  
  # Ensure formative_regime_change is numeric
  data$formative_regime_change <- as.numeric(data$formative_regime_change)
  
  # Control vs Treatment indicators
  control_group <- data$formative_regime_change == 0
  treatment_group <- data$formative_regime_change == 1
  
  # Check that we have both groups
  if(sum(control_group) == 0) stop("No control group (formative_regime_change = 0) found")
  if(sum(treatment_group) == 0) stop("No treatment group (formative_regime_change = 1) found")
  
  # For each variable, perform t-test and extract statistics
  for(var in vars) {
    # Extract values for control and treatment
    control_vals <- data[[var]][control_group]
    treat_vals <- data[[var]][treatment_group]
    
    # Remove NA values
    control_vals <- control_vals[!is.na(control_vals)]
    treat_vals <- treat_vals[!is.na(treat_vals)]
    
    # Calculate means and SDs
    control_mean <- mean(control_vals)
    control_sd <- sd(control_vals)
    treat_mean <- mean(treat_vals)
    treat_sd <- sd(treat_vals)
    
    # Perform t-test
    t_test <- t.test(treat_vals, control_vals)
    diff <- treat_mean - control_mean
    p_value <- t_test$p.value
    
    # Add significance stars
    stars <- ifelse(p_value < 0.001, "***",
                   ifelse(p_value < 0.01, "**",
                         ifelse(p_value < 0.05, "*",
                               ifelse(p_value < 0.1, "+", ""))))
    
    # Format the results
    control_text <- sprintf("%.3f\n(%.3f)", control_mean, control_sd)
    treat_text <- sprintf("%.3f\n(%.3f)", treat_mean, treat_sd)
    diff_text <- sprintf("%.3f%s\n(%.3f)", diff, stars, t_test$stderr)
    
    # Store results
    results[[var]] <- c(
      Variable = var,
      Control = control_text,
      Treatment = treat_text,
      Difference = diff_text
    )
  }
  
  # Convert to data frame
  table_df <- do.call(rbind, results)
  row.names(table_df) <- NULL
  
  # Create the table using kableExtra for PDF output
  require(kableExtra)
  
  # Clean variable names for display
  display_names <- c(
    trust = "Trust",
    patience = "Patience",
    risktaking = "Risk Taking",
    posrecip = "Positive Reciprocity",
    negrecip = "Negative Reciprocity",
    altruism = "Altruism",
    subj_math_skills = "Subjective Math Skills"
  )
  
  # Replace variable names with display names
  table_df[,1] <- ifelse(table_df[,1] %in% names(display_names), 
                        display_names[table_df[,1]], 
                        table_df[,1])
  
  # Create styled table (without caption - handle in document)
  kbl(table_df, 
      booktabs = TRUE,
      col.names = c("", "Control Group", "Treatment Group", "Difference"),
      align = c("l", "c", "c", "c"),
      escape = TRUE) %>%
    kable_styling() %>%
    add_header_above(c(" " = 1, "Mean (SD)" = 3)) %>%
    column_spec(1, width = "2.5cm") %>%
    column_spec(2:4, width = "2cm") %>%
    footnote(general = "Standard deviations in parentheses. Standard errors for differences in parentheses.",
             symbol = c("*** p<0.001, ** p<0.01, * p<0.05, + p<0.1"),
             threeparttable = TRUE,
             general_title = "Note:",
             escape = TRUE)
}
```

```{r}
#| echo: false
#| label: tbl-preferences
#| warning: false
#| results: 'asis'
#| tbl-cap: "Economic Preferences Comparison by Group"
#| tbl-pos: "H"
# Create the table
preferences_vars <- c("trust", "patience", "risktaking", "posrecip", "negrecip", "altruism")
preferences_table <- create_ttest_table(final_data_gdp, preferences_vars)

# Render the table
preferences_table
```

```{r}
#| label: tbl-controls
#| eval: false
#| warning: false
#| include: false
#| results: asis
#| tbl-cap: Control Variables Comparison by Group
#| tbl-pos: H
# Create the table
controls_vars <- c("age", "gdppc_2012", "subj_math_skills")
controls_table <- create_ttest_table(final_data_gdp, controls_vars)

# Render the table
controls_table
```

# Mechanism

Our core hypothesis is that political regime changes during formative years (ages 3--18) have lasting impacts on non-cognitive preferences. This aligns with psychosocial development theory [@balasundaram_human_2025; @detlefsen_are_2024].

We cannot test the mechanism directly, but we test its implications through a Two-way fixed effects (2FE):

-   Individuals are assigned to treatment if they experienced a regime change during their formative years.
-   All individuals are surveyed in the same period (2012--2013), minimizing confounding from external shocks.

We apply a **two-way fixed effects (TWFE)** approach, including region and birth cohort fixed effects. This allows us to account for both: - unobserved heterogeneity across regions, and - time-specific effects common to individuals born in the same period.

Causal interpretation is valid under the **conditional independence assumption**, which implies that---after controlling for observable covariates and fixed effects---the treatment assignment is as good as random.

Possible estimators: [@callaway_difference-differences_2024], [@borusyak_revisiting_2024], [@de_chaisemartin_two-way_2020]. In our current setting, the treatment assignment happens just once.

# Source of Bias and Identification

Our identification strategy leverages variation in individuals' exposure to regime change during their formative years (ages 3--18). We estimate the **Average Treatment Effect on the Treated (ATT)** using a two-way fixed effects (TWFE) framework, which includes region and birth cohort fixed effects.

This specification helps address unobserved heterogeneity across space and time. Causal inference relies on the **conditional independence assumption**---i.e., after controlling for observed covariates and fixed effects, regime change exposure is as good as random.

### Potential Biases and Mitigation

-   **Endogeneity of regime change exposure:**\
    Exposure is not randomly assigned. However, we exploit variation in **birth year within regions**, similar to the quasi-experimental logic in [@fuchs-schundeln_endogeneity_2015].

-   **Time-invariant region-level confounding:**\
    Region fixed effects absorb structural differences across space, such as geography, long-term culture or historical institutions.

-   **Common cohort shocks:**\
    Birth cohort fixed effects capture common temporal shocks (e.g., global crises) that may affect preferences across all individuals born in a given year.

-   **Measurement of regime change:**\
    Regime change is defined as a statistically significant change in LDI over a 10-year window, following [@luhrmann_autocratization_2020], with a stricter threshold of 0.1 applied to capture more sudden transitions.

-   **Residual selection bias:**\
    Countries or regions with more frequent regime changes may also differ economically or politically (e.g., lower GDP per capita). Including controls such as average GDP PC during formative years of individuals along with fixed effects, helps mitigate this concern.

# Formulas

The equation of our main baseline model is as following:

$$
Y_{ict} = \alpha + \beta \cdot \text{RegimeChange}_{ic} + \gamma_c + \delta_t + X_{ict} \cdot \text{(Controls)} + \varepsilon_{ict}
$$ Where:

-   **Outcome:** $Y_{ict}$: Economic preference outcome (trust, patience, etc.) for **individual** ( i ), **region** ( r ), **birth cohort** ( t )

-   **Treatment:** $\text{RegimeChange}_{ic}$: Binary treatment indicator (1 if experienced democratization/autocratization during formative years, 0 otherwise)

-   $\delta_t$ : Birth cohort fixed effects

-   $\gamma_r$ : Region fixed effects

-   $X_{ict}$ : **Controls** including:

    -   Average V-Dem democracy index score experienced during formative years (ages 3-18)

    -   Average GDPPC of the country of residence during formative years (ages 3-18)

    -   Subjective math skills

    We acknowledge that including the "Subjective math skills" as a part of the control variables might be problematic, because of the reverse causality problem.

# Identified Types of Effects

We estimate the **Average Treatment Effect on the Treated (ATT)** using a two-way fixed effects (TWFE) model. This allows us to identify the effect of regime change exposure on individuals who actually experienced such changes during their formative years (ages 3--18).

### Identification Assumptions

The TWFE estimator provides valid causal inference under the **conditional independence assumption**. In our context, this assumption implies:

-   Conditional on observed covariates, region fixed effects, and birth cohort fixed effects, exposure to regime change is **as good as random**.
-   There are **no unobserved confounders** that vary simultaneously with both treatment assignment and the preference outcomes.
-   Individuals not exposed to regime change during formative years serve as a valid control group for those who were exposed.

We control for key observed characteristics (e.g., GDP per capita, subjective math skills) and absorb unobserved heterogeneity at the region and cohort levels to satisfy this assumption as closely as possible.

# Questions for meeting on 12.3.

-   After running the basic regression -\> effect of formative_regime_change significant but low R2

-   Regarding the methodology as described in "Setting": We are capturing democratization/autocratization as a regime change. But maybe not all democratization/autocratization process must necessarily mean the regime has changed? Should we consider another criterion which at the end evaluates if the value of LDI is: above 0.5 -\> liberal democracy, below 0.5 -\> autocracy?

-   Treatment assignment: With the current approach, there is a lot of observations who, according to our function, experience regime change at age 0. It is because the function evaluates just the first democratization/autocratization occurence. But I keep thinking there could be more such episodes? Should we aim to capture also other episodes OR change the confidence intervals threshold? This problem also connects to the next question:

-   How do we calculate the years that have passed since the new regime change? For example, maybe a person changed regimes twice in his formative years? How do we calculate this? Or just we count from the first change?

-   Adding interaction term $TimeToTreatment_{c,t,k}$ as an indicator for being $k$ **periods away from the treatment** (regime change) in country $c$ at time $t$?

average democratization between 16-25 -\> treatment just when its sudden germany how much change in vdem in 1990? capture sudden changes rather than trends

average vdem score the individual experience in formative years -\> maybe more feasible democratization/autocratization dummy as a treatment

regress econ preferences on average vdem scores experienced in youth average GDP between birthy-year and year adult

as control age effects and cohort effects -\> use just one we cant use both pseudo panel -\> event study -\> rather go with cross sections -\> vdem index in formative years influences preferences country

conditial independence assumption -\> still can be causal

math skills can also be an outcome

change just to regime_change

simple OLS just cross section -\> if we trust conditional independence assumption -\> how to make use of different cohorts in the dataset -\> instrument?

focus on average vdem index in formative years staggered DiD: years relative to treatment

# References
