---
title: "Regime Changes and Economic Preferences: Global Evidence"
subtitle: "Milestone 5: Robustness"
author: "Andrea Češková and Elvin Mammadov"
format: 
  pdf:
    reference-location: document
number-sections: true
toc: true
crossref:
  fig-title: "Figure"
  tbl-title: "Table"
  eq-title: "Equation"
  lot-title: "List of figures"
bibliography: references.bib
link-citations: true
---

```{r}
#| message: false
#| warning: false
#| include: false
library(pacman)

p_load(haven,
       here,
       stargazer,
       summarytools,
       readxl,
       dplyr,
       lubridate,
       ggplot2,
       vdemdata,
       summarytools,
       devtools,
       zoo,
       tidyverse,
       fixest,
       dplyr,
       tidyr,
       modelsummary,
       tinytable,
       here,
       countrycode,
       gt,
       knitr,
       kableExtra
       )
```


```{r}
#| include: false
conditional_balance <- readRDS(here("Output", "Tables", "conditional_balance_table.rds"))
democratization_balance <- readRDS(here("Output", "Tables", "balance_table_democratization_vs_autocratization.rds"))
overall_balance <- readRDS(here("Output", "Tables", "balance_table_overall.rds"))
sample_composition <- readRDS(here("Output", "Tables", "sample_composition.rds"))
clustering_results <- readRDS(here("Output", "Robustness", "clustering_results.rds"))
country_summary <- readRDS(here("Input", "clean", "country_summary.rds"))

# Regime changes visualization
# Polity5
all_countries_plot_polity <- readRDS(here("Output", "figures", "all_countries_plot_polity.rds"))
# RWI
all_countries_plot_RWI <- readRDS(here("Output", "figures", "all_countries_plot_RWI.rds"))
```

```{r}
#| message: false
#| warning: false
#| include: false
demo_summary <- clustering_results$demo_summary
auto_summary <- clustering_results$auto_summary
```

# Sample composition

Our final analytical sample consists of `r sum(sample_composition$n)` individuals across three groups:

```{r}
#| echo: false
#| label: tbl-samplecomp
#| tbl-cap: "Sample Composition by Treatment Type"
#| tbl-pos: "h"
create_summary_table <- function() {
  
  # Defining group labels 
  group_labels <- c(
    "0" = "Control (No regime change)",  
    "1" = "Autocratization",              
    "2" = "Democratization"             
  )
  
  # Calculating summary statistics
  summary_table <- country_summary %>%
    group_by(first_formative_type) %>%
    summarise(
      N = sum(n_individuals, na.rm = TRUE),
      .groups = 'drop'
    ) %>%
    mutate(
      Group = case_when(
        first_formative_type == 0 ~ "Control (No regime change)",
        first_formative_type == 1 ~ "Autocratization", 
        first_formative_type == 2 ~ "Democratization",
        TRUE ~ as.character(first_formative_type)
      ),
      Proportion = round(N / sum(N), 2),
      Percentage = round((N / sum(N)) * 100, 1)
    ) %>%
    select(Group, N, Proportion, Percentage) %>%
    arrange(desc(N))  # Sorting by sample size, largest first
  
  return(summary_table)
}

# Generating the table
table_data <- create_summary_table()

# Creating a formatted table
formatted_table <- table_data %>%
  kable(
    col.names = c("Group", "N", "Proportion", "Percentage"),
    align = c("l", "r", "r", "r"),
    format = "latex"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE,
    position = "center"
  ) %>%
  add_footnote(
    "Distribution of observations across control and treatment groups.",
    notation = "none"
  )

# Displaying the formatted table
formatted_table


```

# Different specification

## Clustering of standard errors

Our baseline specification clusters standard errors at the country level, reflecting our assumption that the primary source of correlation in errors stems from shared country-specific factors such as cultural norms, institutional quality, and historical experiences. To assess the robustness of our inference, we examine how our results change under alternative clustering assumptions.

In this chapter we present sensitivity analysis across four clustering methods: **country-level clustering (our baseline)**, **two-way clustering by country and birth cohort**, **regional clustering**, and **birth cohort clustering**. We chose to analyse only our baseline specification. @tbl-cluster_democ shows the results for base specification: democratization and @tbl-cluster_autoc for base specification: autocratization.

```{r}
#| message: false
#| warning: false
#| include: false

# Function to create professional clustering tables for PDF
create_clustering_table_pdf <- function(summary_data, treatment_type) {
  
  summary_data$preference <- c("Patience", "Risk taking", "Positive Reciprocity", "Negative Reciprocity", "Altruism", "Trust")
  # Clean up column names to match your report style
  colnames(summary_data) <- c("", "Country", "Two-way", "Region", "Cohort")
  
  # Create the table for PDF/LaTeX output
  table <- summary_data %>%
    kable(format = "latex", 
          align = c("l", "c", "c", "c", "c"),
          escape = FALSE) %>%
    kable_styling(
      latex_options = c("striped", "scale_down"),
      full_width = FALSE,
      position = "center",
      font_size = 10
    ) %>%
    row_spec(0, bold = TRUE) %>%
    column_spec(1, width = "4cm", bold = FALSE) %>%
    column_spec(2:5, width = "2cm") %>%
    add_header_above(c(" " = 1, "Clustering Method" = 4), bold = TRUE) %>%
    footnote(
      general = "Standard errors clustered as indicated in column headers. All models include region and birth year fixed effects.",
      general_title = "Note:",
      footnote_as_chunk = TRUE,
      threeparttable = TRUE
    )
  
  return(table)
}

```

The patience affect in @tbl-cluster_democ remains stable across all clustering methods, with standard errors ranging from 0.0207 to 0.0262, maintaining significance at the 5% level or better. This consistency provides **strong evidence** that our main finding---**experiencing democratization during formative years increases patience**---is not just because of our choice of one clustering method.

The similarity between standard errors under country-level (0.025) and two-way clustering (0.024) indicates minimal temporal correlation within countries. Regional clustering yields slightly larger standard errors, as expected given the smaller number of clusters, while cohort clustering produces the smallest standard errors.

Notably, the **altruism coefficient** exhibits sensitivity to the clustering method. While the point estimate remains stable at -0.078, it is insignificant under country clustering but becomes **highly significant (p \< 0.001)** under **cohort clustering**. This pattern suggests that altruism preferences exhibit **stronger within-country correlation than within-cohort correlation** - in other words, individuals within the same country are more similar to each other in their altruism preferences than individuals born in the same year globally. This likely reflects the importance of country-specific cultural transmission mechanisms in shaping other-regarding preferences.

```{r}
#| echo: false
#| label: tbl-cluster_democ
#| tbl-cap: "Sensitivity to Clustering Approaches - Democratization Effects"
#| tbl-pos: "h"
#| fig-pos: "h"
demo_table_pdf <- create_clustering_table_pdf(demo_summary, 
                                              "democratization")

demo_table_pdf
```

Results in @tbl-cluster-autoc how generally weaker effects across all clustering methods, consistent with our main findings. The marginal significance of positive reciprocity is relatively stable across specifications, while other effects remain statistically insignificant regardless of clustering approach.

```{r}
#| echo: false
#| label: tbl-cluster_autoc
#| tbl-cap: "Sensitivity to Clustering Approaches - Autocratization Effects"
#| tbl-pos: "h"
#| fig-pos: "h"
auto_table_pdf <- create_clustering_table_pdf(auto_summary, 
                                              "autocratization")
auto_table_pdf
```

Overall, these robustness checks strengthen confidence in our core findings while highlighting which results are sensitive to different clustering approaches. The **persistence of the democratization-patience relationship** across all specifications underscores its robustness as our primary empirical contribution.

# Different Datasets

## V-dem: Regime of the World Index

V2xregime value from VDem data, created by the Varieties of Democracy (V-Dem) Project, classifies countries based on their regime type using a four-category system that reflects the degree of democracy and autocracy. The classification is as follows:

**0 --- Closed Autocracies**: Countries where citizens have no meaningful opportunity to participate in elections and political power is highly concentrated.

**1 --- Electoral Autocracies**: Countries that hold de jure multiparty elections, but these are neither free nor fair and core democratic rights are restricted.

**2 --- Electoral Democracies**: Countries where elections are relatively free and fair, but democratic institutions (like judicial independence or media freedom) may be weak or partially undermined.

**3 --- Liberal Democracies**: Countries that not only hold free and fair elections but also uphold civil liberties, the rule of law and strong checks and balances on executive power.

These categories are based on V-Dem's extensive dataset, which includes expert-coded indicators on political institutions, rights and governance. The regime change included only from Autocracy to Democracy or vice verca.

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 16
#| fig-height: 12
#| fig-cap: "Regime Changes by Country and Continent (Post-1920)"
#| label: fig-regime-changes-RWI

all_countries_plot_RWI
```

### Base Specification with Heterogenous Effects

```{r}
#| message: false
#| warning: false
#| include: false

 regime_change_lde <- readRDS(here("Output", "Models", "regime_change_models.rds"))

 
 regime_change_models_v2xregime <- readRDS(here("Output" , "Models" ,"v2xregime results", "regime_change_models_v2xregime.rds"))
```

### Fixed effects: Regime change comparison

```{r}
#| echo: false
#| label: tbl-regime_regime_lde
#| tbl-cap: "The results of regime and LDE data based models (Regime change comparison)"
#| tbl-pos: "h"
#| fig-pos: "h"
# Define preference names for titles
preference_names <- c("Patience", "Risk taking", "Positive reciprocity", 
                      "Negative reciprocity", "Altruism", "Trust")

# Generate a comparison table for each preference
comparison_tables <- lapply(seq_along(preference_names), function(i) {
  msummary(
    list(
      "Regime-based model" = regime_change_models_v2xregime[[i]],
      "LDE-based model"    = regime_change_lde[[i]]
      
    ),
    stars = TRUE,
    coef_rename = c(
      "regime_change" = "Regime Change Indicator"
    ),
    title = paste("Preference:", preference_names[i]),
    notes = "Standard errors clustered at isocode (country) level. Fixed effects and controls included.",
    output = "gt"
  )
})

# Display each table one after another
for (tbl in comparison_tables) {
  print(tbl)
}
```

LDE-based model: Formative regime change has a strong positive and highly significant effect on patience and strongly positive effect on negative reciprocity. But in regime model there is no significant effect

###hetero models comparision Combined: autocratization and democratization

```{r}
#| message: false
#| warning: false
#| include: false

# Define preference names for titles
hetero_models_v2 <- readRDS(here("Output", "Models", "v2xregime results", "hetero_models_v2xregime.rds"))
hetero_models_democ_v2 <- readRDS(here("Output", "Models", "v2xregime results", "hetero_models_democ_v2xregime.rds"))
hetero_models_autoc_v2 <- readRDS(here("Output", "Models", "v2xregime results", "hetero_models_autoc_v2xregime.rds"))


# LDE model
lde_combined <- readRDS(here("Output", "Models", "hetero_models.rds"))
lde_democ    <- readRDS(here("Output", "Models","hetero_models_democ.rds"))
lde_autoc    <- readRDS(here("Output", "Models","hetero_models_autoc.rds"))

```

```{r}
#| echo: false
#| label: tbl-fe_regime_lde_hetero
#| tbl-cap: "The results of regime and LDE data based models (hetero models comparision Combined)"
#| tbl-pos: "h"
#| fig-pos: "h"
comparison_tables <- lapply(seq_along(preference_names), function(i) {
  msummary(
    list(
      "Regime: Combined"        = hetero_models_v2[[i]],
      "Regime: Democratization" = hetero_models_democ_v2[[i]],
      "Regime: Autocratization" = hetero_models_autoc_v2[[i]],
      "LDE: Combined"           = lde_combined[[i]],
      "LDE: Democratization"    = lde_democ[[i]],
      "LDE: Autocratization"    = lde_autoc[[i]]
    ),
    stars = TRUE,
    coef_rename = c(
      "democratization" = "Democratization",
      "autocratization" = "Autocratization"
    ),
    title = preference_names[i],
    notes = "Standard errors clustered at isocode (country) level. Fixed effects for region and birth year included.",
    output = "gt"
  )
})


for (tbl in comparison_tables) {
  print(tbl)
}
```

In regime models, Autocratization has strongly negative effect on risk taking and trust of individiuals, also it has also effect on negative reciprocity, but this one is strongly significant.

### Specification with Controls

```{r}
#| message: false
#| warning: false
#| include: false

# Read models first
hetero_models_controls_v2 <- readRDS(here("Output", "Models", "v2xregime results", "hetero_models_controls_v2xregime.rds"))
hetero_models_democ_controls_v2 <- readRDS(here("Output", "Models", "v2xregime results", "hetero_models_democ_controls_v2xregime.rds"))
hetero_models_autoc_controls_v2 <- readRDS(here("Output", "Models", "v2xregime results", "hetero_models_autoc_controls_v2xregime.rds"))




# LDE-based model
lde_controls_combined <- readRDS(here("Output", "Models","hetero_models_controls.rds"))
lde_controls_democ    <- readRDS(here("Output", "Models","hetero_models_democ_controls.rds"))
lde_controls_autoc    <- readRDS(here("Output", "Models","hetero_models_autoc_controls.rds"))
```

```{r}
#| echo: false
#| label: tbl-fe_regime_lde_controls
#| tbl-cap: "The results of regime and LDE data based models (Specification with Controls)"
#| tbl-pos: "h"
#| fig-pos: "h"
# creating list of table
comparison_tables_controls <- lapply(seq_along(preference_names), function(i) {
  msummary(
    list(
      "Regime: Combined"        = hetero_models_controls_v2[[i]],
      "Regime: Democratization" = hetero_models_democ_controls_v2[[i]],
      "Regime: Autocratization" = hetero_models_autoc_controls_v2[[i]],
      "LDE: Combined"           = lde_controls_combined[[i]],
      "LDE: Democratization"    = lde_controls_democ[[i]],
      "LDE: Autocratization"    = lde_controls_autoc[[i]]
    ),
    stars = TRUE,
    coef_rename = c(
      "democratization" = "Democratization",
      "autocratization" = "Autocratization",
      "avg_gdppc_formative" = "Log of Average GDP per Capita during formative years",
      "recession_formative" = "Recession experienced during formative years"
    ),
    title = preference_names[i],
    notes = "Standard errors clustered at isocode (country) level. All models include region and birth year fixed effects.",
    output = "gt"
  )
})

# Show the results
for (tbl in comparison_tables_controls) {
  print(tbl)
}

```

Patience is positively and strongly associated with the log of average GDP per capita during formative years across all models. This suggests that individuals who grew up in wealthier economic environments tend to be more patient. Risk taking significantly decreases following autocratization in regime-based models. Negative reciprocity significantly declines after autocratization in regime models , indicating that such political transitions may suppress retaliatory economic behaviors. Trust decreases substantially in response to autocratization in regime-based models, highlighting a strong erosion of interpersonal or institutional trust in individuals exposed to such regime shifts.

## Polity dataset

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 16
#| fig-height: 12
#| fig-cap: "Regime Changes by Country and Continent (Post-1920)"
#| label: fig-regime-changes-polity

all_countries_plot_polity
```

```{r}
#| message: false
#| warning: false
#| include: false

 regime_change_lde <- readRDS(here("Output", "Models", "regime_change_models.rds"))

 
 regime_change_models_polity <- readRDS(here("Output" , "Models" ,"polity results", "regime_change_models_polity.rds"))
```

Coverage: 160+ countries, typically from 1800–2023

Main variable: polity2, a composite score ranging from –10 (full autocracy) to +10 (full democracy) threshold for polity2 : polity2 \>= +6 is determined as Democratization, polity2 \< +6 Autocratization

Tracks regime changes, authority characteristics and durability of political institutions

```{r}
#| echo: false
#| label: tbl-fe_polity_lde_regime
#| tbl-cap: "The results of polity and LDE data based models (Regime change)"
#| tbl-pos: "h"
#| fig-pos: "h"
# Define preference names for titles
preference_names <- c("Patience", "Risk taking", "Positive reciprocity", 
                      "Negative reciprocity", "Altruism", "Trust")

# Generate a comparison table for each preference
comparison_tables <- lapply(seq_along(preference_names), function(i) {
  msummary(
    list(
      "Polity-based model" = regime_change_models_polity[[i]],
      "LDE-based model"    = regime_change_lde[[i]]
    ),
    stars = TRUE,
    coef_rename = c(
      "regime_change" = "Regime Change Indicator"
    ),
    title = paste("Preference:", preference_names[i]),
    notes = "Standard errors clustered at isocode (country) level. Fixed effects and controls included.",
    output = "gt"
  )
})

# Display each table one after another
for (tbl in comparison_tables) {
  print(tbl)
}

```

Compare to the LDE model and regime based model , in Polity-based model, formative regime change has strongly negative effect on altuism.

###hetero models comparision Combined: autocratization and democratization

```{r}
#| message: false
#| warning: false
#| include: false

# Load models from the "polity results" folder
hetero_models_polity <- readRDS(here("Output", "Models", "polity results", "hetero_models_polity.rds"))
hetero_models_democ_polity <- readRDS(here("Output", "Models", "polity results", "hetero_models_democ_polity.rds"))
hetero_models_autoc_polity <- readRDS(here("Output", "Models", "polity results", "hetero_models_autoc_polity.rds"))

```

```{r}
#| echo: false
#| label: tbl-fe_polity_lde_hetero
#| tbl-cap: "The results of polity and LDE data based models (hetero models comparision Combined)"
#| tbl-pos: "h"
#| fig-pos: "h"
# Comparison table: Polity-based models vs. LDE-based models
comparison_tables <- lapply(seq_along(preference_names), function(i) {
  msummary(
    list(
      "Polity: Combined"        = hetero_models_polity[[i]],
      "Polity: Democratization" = hetero_models_democ_polity[[i]],
      "Polity: Autocratization" = hetero_models_autoc_polity[[i]],
      "LDE: Combined"           = lde_combined[[i]],
      "LDE: Democratization"    = lde_democ[[i]],
      "LDE: Autocratization"    = lde_autoc[[i]]
    ),
    stars = TRUE,
    coef_rename = c(
      "democratization" = "Democratization",
      "autocratization" = "Autocratization"
    ),
    title = preference_names[i],
    notes = "Standard errors clustered at isocode (country) level. All models include region and birth year fixed effects.",
    output = "gt"
  )
})

# Print each comparison table
for (tbl in comparison_tables) {
  print(tbl)
}
```

### Specification with Controls

```{r}
#| message: false
#| warning: false
#| include: false

# Read polity-based models with controls
hetero_models_controls_polity <- readRDS(here("Output", "Models", "polity results", "hetero_models_controls_polity.rds"))
hetero_models_democ_controls_polity <- readRDS(here("Output", "Models", "polity results", "hetero_models_democ_controls_polity.rds"))
hetero_models_autoc_controls_polity <- readRDS(here("Output", "Models", "polity results", "hetero_models_autoc_controls_polity.rds"))

```

```{r}
#| echo: false
#| label: tbl-fe_polity_lde_controls
#| tbl-cap: "The results of polity and LDE data based models (Specification with Controls)"
#| tbl-pos: "h"
#| fig-pos: "h"
# Create comparison tables with controls: Polity vs. LDE
comparison_tables_controls <- lapply(seq_along(preference_names), function(i) {
  msummary(
    list(
      "Polity: Combined"        = hetero_models_controls_polity[[i]],
      "Polity: Democratization" = hetero_models_democ_controls_polity[[i]],
      "Polity: Autocratization" = hetero_models_autoc_controls_polity[[i]],
      "LDE: Combined"           = lde_controls_combined[[i]],
      "LDE: Democratization"    = lde_controls_democ[[i]],
      "LDE: Autocratization"    = lde_controls_autoc[[i]]
    ),
    stars = TRUE,
    coef_rename = c(
      "democratization" = "Democratization",
      "autocratization" = "Autocratization",
      "avg_gdppc_formative" = "Log of Average GDP per Capita during formative years",
      "recession_formative" = "Recession experienced during formative years"
    ),
    title = preference_names[i],
    notes = "Standard errors clustered at isocode (country) level. All models include region and birth year fixed effects.",
    output = "gt"
  )
})

# Display each table
for (tbl in comparison_tables_controls) {
  print(tbl)
}

```

But in all 3 different data based models, we can see that Log of Average GDP per Capita has strongly positive effect on Patience. This coefficient from Polity based model is way more than rest 2 different data based models.
