---
title: "Data wragling"
date: "2025-04-20"
output: html_document 
---

```{r Loading the Data}

library(pacman)
p_load(haven,
       here,
       stargazer,
       summarytools,
       readxl,
       dplyr,
       lubridate,
       ggplot2,
       vdemdata,
       summarytools,
       devtools,
       zoo,
       tidyverse)


source(here("R", "Script", "01_data_loading.R"))

```

# Data wrangling

Questions for meeting on 29.04. :

-   **How to decide on a threshold of assigning *generate_regime_change*?** So far, we have been working with liberal democracy index from V-Dem, but there are more options -\> should we aggregate them into one?

    *answer whether its a sudden change or linear -\> which are better suited?*

*Restrict just to one schock: autoritative -\> democratic*

-   **Separate model for each preference / try to combine them?** *check for literature- maybe choose 2 based on this, if we use all of them, Lessmann will ask about multiple hypotheses testing :)*

*time, age, cohort fixed effects*

*age/cohort effects could be problematic*

*net out age effect from economic preferences -\>whats the effect of risk aversion? (next step)*

*if not DiD: mean comparisons -\> give up on causal effects -\> more about living under regime than regime changes -\>already dome evidence on this*

```{r}
source(here("R", "Script", "02_filtering.R"))
```

## Treatment calculation

### First step: identifying regime changes on country level

```{r regime changes periods on country level}
source(here("R", "Script", "03_regime_changes_function.R"))
```

### Second step: connecting with GPS data

```{r}
source(here("R", "Script", "04_merging_GPS_vdem.R"))
```

## Contaminated control group

The approach above assigns following individuals to be part of the control group:

-   **True controls:** Individuals from countries that never experienced any regime changes

-   **Contaminated controls:** Individuals from countries that did experience regime changes, but these individuals happen to be outside of their formative years when it occured -\> problematic, because they might have indirect exposure effect, they still lived through regime changes as adults outside of formative years. The following part of the code removes these individuals from out dataset.

```{r exploring the both groups}
source(here("R", "Script", "05_exploring_both_groups.R"))
```

## Calculating controls

The following code takes a bit longer to generate because of a nested loop structure. It calculates the following controls:

-   Average GDP in the country of residence during formative years of the individual (3-18 years)

-   Average LDI (Liberal Democracy Index) i nthe country of residence during formative years of the individual (3-18 years)

```{r average gdppc and democracy index}

source(here("R", "Script", "06_average_gdppc_index.R"))
```

During the robustness check, we explored our options how to add more controls. In the following part of the code we include:

-   Dummy for a recession during formative years

The recession is calculated based on country-specific volatility

```{r recession dummy}

source(here("R", "Script", "07_recession_dummy.R"))
```

## Saving the datasets in RDS format for later use

```{r}
# Saving the final datasets in rds format

# data all osbervations
saveRDS(final_data, here("Input","clean", "final_data.rds"))

# data filtered observations with controls
saveRDS(final_data_gdp, here("Input","clean", "final_data_gdp.rds"))

# data filtered observations with controls
saveRDS(final_data_clean, here("Input","clean", "final_data_clean.rds"))

# country summary data for tables in Milestone 3 report
saveRDS(country_summary, here("Input","clean", "country_summary.rds"))

```
